Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-03-20 paper.qmd
Updated Figure 12 with the upated code

@@

generate_error_plots_one_c_shaped()
```

>>>>

```{r}
#| label: model-one-dens-clust

one_c_shaped_data <- read_rds(here::here("data/one_c_shaped_dens_structure/one_c_shaped_dens_data.rds"))
one_c_shaped_data <- one_c_shaped_data |>
  mutate(ID = row_number())

tsne_one_c_shaped <- read_rds(file = "data/one_c_shaped_dens_structure/one_c_shaped_dens_structure_tsne_perplexity_52.rds")

num_bins_x_one_c_shaped <- 20

algo_obj_dens_clust <- gen_nldr_vis_algo_obj(
  high_d_data = one_c_shaped_data, 
  nldr_data = tsne_one_c_shaped, 
  num_x_bins = num_bins_x_one_c_shaped,
  col_start = "x")

tsne_one_c_shaped_scaled <- algo_obj_dens_clust$nldr_scaled
distance_dens_clust <- algo_obj_dens_clust$distance_df
tr_from_to_df_dens_clust <- algo_obj_dens_clust$tr_from_to_df
benchmark_dens_clust <- algo_obj_dens_clust$benchmark
df_bin_centroids_dens_clust <- algo_obj_dens_clust$df_bin_centroids
df_bin_dens_clust <- algo_obj_dens_clust$df_bin
hex_grid_with_counts_dens_clust <- algo_obj_dens_clust$hex_grid_with_counts
```

```{r}
#| label: error-model-one-dens-clust
## Compute error
error_df_one_curvy_abs <- augment(
  df_bin_centroids = df_bin_centroids_dens_clust,
  df_bin = df_bin_dens_clust,
  training_data = one_c_shaped_data,
  newdata = NULL,
  type_NLDR = "tSNE",
  col_start = "x")

error_df_one_curvy_abs <- error_df_one_curvy_abs |>
  bind_cols(tsne_one_c_shaped_scaled |>
              select(-ID))

error_df_one_curvy_abs <- error_df_one_curvy_abs |>
  mutate(sqrt_row_wise_total_error = sqrt(row_wise_total_error))

# Compute density
density_data <- density(error_df_one_curvy_abs$sqrt_row_wise_total_error)
density_df <- data.frame(x = density_data$x, y = density_data$y)

# Add density values to the original dataset
error_df_one_curvy_abs <- error_df_one_curvy_abs %>%
  mutate(density = approx(density_df$x, density_df$y, xout = sqrt_row_wise_total_error)$y)

error_df_one_curvy_abs <- error_df_one_curvy_abs |>
  mutate(error_cat_n = if_else(tSNE1 <= 0.25 & tSNE2 <= 0.25, "high", "low")) |> ## high_error points
  mutate(error_cat_n2 = if_else(tSNE2 >= 1.2, "high", "low")) ## corner points

error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs, aes(x = sqrt_row_wise_total_error, y = density,
                                                                         colour = factor(error_cat_n,
                                                                                         levels = c("high", "low")))) +
  geom_point() +
  scale_color_manual(values=c('#800026', '#d9d9d9')) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("b1",
                      position = c(0.95, 0.9),
                      cex = 1.5) +
  theme_bw() +
  theme(
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 15)
  )

error_plot_one_curvy_hist <- ggplot(error_df_one_curvy_abs,
                                    aes(x = sqrt_row_wise_total_error, y = density,
                                        colour = factor(error_cat_n2,
                                                        levels = c("high", "low")))) +
  geom_point() +
  scale_color_manual(values=c('#800026', '#d9d9d9')) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("a1",
                      position = c(0.95, 0.9),
                      cex = 1.5) +
  theme_bw() +
  theme(
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 15)
  )
```

```{r}
#| label: link-model-one-dens-clust
plot_tsne_dens <- error_df_one_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = factor(error_cat_n2,
                             levels = c("high", "low")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#800026', '#d9d9d9')) +
  theme(
    aspect.ratio = 1
  ) +
  interior_annotation("a2",
                      position = c(0.08, 0.9),
                      cex = 1.5)

plot_tsne_dens_selected <- error_df_one_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = factor(error_cat_n,
                             levels = c("high", "low")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#800026', '#d9d9d9')) +
  theme(
    aspect.ratio = 1
  ) +
  interior_annotation("b2",
                      position = c(0.08, 0.9),
                      cex = 1.5)

distance_df_small_edges_c_shaped_structure <- distance_dens_clust |>
  filter(distance < benchmark_dens_clust)

tr_from_to_df_dens_clust <- inner_join(
  tr_from_to_df_dens_clust, distance_df_small_edges_c_shaped_structure,
  by = c("from", "to"))

```{r}
#| label: proj-one-dens-clust

# Apply the scaling

data_c_shaped <- one_c_shaped_data |>
  select(-ID) |>
  mutate(type = "data")

df_bin_dens_clust <- df_bin_dens_clust |>
  select(-hb_id) |>
  mutate(type = "model")

df_model_data <- bind_rows(data_c_shaped, df_bin_dens_clust)
scaled_c_shaped <- scale_data_manual(df_model_data, "type") |>
  as_tibble()

scaled_c_shaped_data <- scaled_c_shaped |>
  filter(type == "data") |>
  select(-type)

scaled_c_shaped_data_model <- scaled_c_shaped |>
  filter(type == "model") |>
  select(-type)

## First projection
projection <- cbind(
  c(0.05096,0.15399,0.19736,0.05110),
  c(0.14608,-0.16929,0.11291,-0.07160))

proj_obj1 <- get_projection(projection = projection,
                            proj_scale = 1,
                            scaled_data = scaled_c_shaped_data,
                            scaled_data_model = scaled_c_shaped_data_model,
                            distance_df_small_edges = distance_df_small_edges_c_shaped_structure,
                            axis_param = list(limits = 0.35,
                                              axis_scaled = 2,
                                              axis_pos_x = -0.28,
                                              axis_pos_y = -0.28,
                                              threshold = 0.017))

projected_df <- proj_obj1$projected_df
model_df <- proj_obj1$model_df
axes2 <- proj_obj1$axes
circle <- proj_obj1$circle

## To add error category
projected_df <- projected_df |>
  mutate(cluster = error_df_one_curvy_abs$error_cat_n2)

five_c_shaped_proj_tsne_model1 <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes2, 
    circle = circle, 
    point_param = c(1.5, 0.5), # size, alpha, color
    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
    plot_limits = c(-0.35, 0.35), 
    title = "a3", 
    cex = 1.5,
    position = c(0.08, 0.9),
    axis_text_size = 4, 
    is_category = TRUE) +
    scale_color_manual(values = c('#800026', '#d9d9d9')) +
    theme(
        legend.position = "none"
    )

projected_df <- projected_df |>
  select(-cluster) |>
  mutate(cluster = error_df_one_curvy_abs$error_cat_n)

five_c_shaped_proj_tsne_model1_selected <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes2, 
    circle = circle, 
    point_param = c(1.5, 0.5), # size, alpha, color
    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
    plot_limits = c(-0.35, 0.35), 
    title = "a3", 
    cex = 1.5,
    position = c(0.08, 0.9),
    axis_text_size = 4, 
    is_category = TRUE) +
    scale_color_manual(values = c('#800026', '#d9d9d9')) +
    theme(
        legend.position = "none"
    )
```

@@

error_plot_one_curvy_hist + plot_tsne_dens + five_c_shaped_proj_tsne_model1 +
  error_plot_one_curvy_hist_selected + plot_tsne_dens_selected + five_c_shaped_proj_tsne_model1_selected +
    plot_layout(guides = "collect", nrow = 2) &
    theme(legend.position='none')
```
